Need Fulfillment During Intergroup Contact: Three Experience Sampling Studies

### 2023-05-30 manuscript.tex
typos

% Abstract
\abstract{
One challenge of modern intergroup contact research has been the question of when and why an interaction is perceived as positive and leads to better intergroup relations. We propose to consider situational psychological needs during everyday intergroup contact. We conducted three extensive longitudinal studies with recent migrants, to capture their interactions with the majority outgroup (total \textit{N} of measurements = 10,297; total \textit{N} participants = 207). The novel psychological needs mechanism is consistently a strong predictor of positive attitudes during intergroup interactions and these positive effects emerged via perceived interaction quality. The situational needs model is robust to alternative models and outperforms Allport's contact conditions. As one of the first studies to test intergroup contact theory using extensive longitudinal data, we offer insight into the mechanisms of positive intergroup contact during real-life interactions and find situational motivations to be a key building block of understanding and addressing positive intergroup interactions.

\noindent\textbf{Public significance statement}: In this paper, we provide evidence that the fulfillment of situational psychological needs during real-life intergroup contacts meaningfully predicts perceived interaction quality and positive outgroup attitudes. Methodologically, this offers testament to the emerging practice of capturing real-life interactions using extensive longitudinal data. Theoretically, our results give weight to motivational fulfillment as a flexible and effective mechanism for understanding positive intergroup contact.
}

\keywords{
    Intergroup Contact, Psychological Needs, Outgroup Attitudes, Interaction Quality, Extensive Longitudinal Data\\
    \textit{Open Science Practices:}
    \noindent \href{https://osf.io/r5e8c?view_only=bfaed2196d49420385ce5fa6506a7765}{\includegraphics[height=\badgeheight]{assets/open-badges-small/registrationplus-color.png}} Preregistration+, 
    \href{https://osf.io/pr9zs/?view_only=1ea47bb646694632a764dead807ef970}{\includegraphics[height=\badgeheight]{assets/open-badges-small/material-color.png}} Open Materials, 
    \href{https://osf.io/pr9zs/?view_only=1ea47bb646694632a764dead807ef970}{\includegraphics[height=\badgeheight]{assets/open-badges-small/data-color.png}} Open Data,\break
    \href{https://osf.io/pr9zs/?view_only=1ea47bb646694632a764dead807ef970}{\includegraphics[height=\badgeheight]{assets/open-badges-small/code-color.png}} Open Code, 
    \href{https://osf.io/pr9zs/?view_only=1ea47bb646694632a764dead807ef970}{\includegraphics[height=\badgeheight]{assets/open-badges-small/supplements-color.png}} Open Supplements
    }

>>>>

% Abstract
\abstract{
One challenge of modern intergroup contact research has been the question of when and why an interaction is perceived as positive and leads to better intergroup relations. We propose to consider the fulfillment of situationally relevant needs during everyday intergroup contact. We conducted three extensive longitudinal studies with recent migrants, to capture their interactions with the majority outgroup (total \textit{N} of measurements = 10,297; total \textit{N} participants = 207). The novel need fulfillment mechanism is consistently a strong predictor of perceived interaction quality as well as positive outgroup attitudes following intergroup interactions. The situational needs model is robust to alternative models and works at least as well as Allport's contact conditions. As one of the first studies to test intergroup contact theory using extensive longitudinal data, we offer insight into the mechanisms of positive intergroup contact during real-life interactions and find situational motivations to be a key building block of understanding and addressing positive intergroup interactions.

\noindent\textbf{Public significance statement}: In this paper, we provide evidence that the fulfillment of situational needs during real-life intergroup contacts meaningfully predicts perceived interaction quality and positive outgroup attitudes. Methodologically, this offers testament to the emerging practice of capturing real-life interactions using extensive longitudinal data. Theoretically, our results give weight to motivational fulfillment as a flexible and effective mechanism for understanding positive intergroup contact.
}

\keywords{
    Intergroup Contact, Need Fulfillment, Outgroup Attitudes, Interaction Quality, Extensive Longitudinal Data\\
    \textit{Open Science Practices:}
    \noindent \href{https://osf.io/r5e8c?view_only=bfaed2196d49420385ce5fa6506a7765}{\includegraphics[height=\badgeheight]{assets/open-badges-small/registrationplus-color.png}} Preregistration+, 
    \href{https://osf.io/pr9zs/?view_only=1ea47bb646694632a764dead807ef970}{\includegraphics[height=\badgeheight]{assets/open-badges-small/material-color.png}} Open Materials, 
    \href{https://osf.io/pr9zs/?view_only=1ea47bb646694632a764dead807ef970}{\includegraphics[height=\badgeheight]{assets/open-badges-small/data-color.png}} Open Data,\break
    \href{https://osf.io/pr9zs/?view_only=1ea47bb646694632a764dead807ef970}{\includegraphics[height=\badgeheight]{assets/open-badges-small/code-color.png}} Open Code, 
    \href{https://osf.io/pr9zs/?view_only=1ea47bb646694632a764dead807ef970}{\includegraphics[height=\badgeheight]{assets/open-badges-small/supplements-color.png}} Open Supplements
    }
